---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ensumr

<!-- badges: start -->
<!-- badges: end -->

The goal of ensumr is to analyse enquate data easily. 

## Installation

You can install the development version of ensumr from [GitHub](https://github.com/) with:

```{r setup, include = FALSE}
devtools::load_all(".")
```


```{r, eval = FALSE}
  # install.packages("remotes")
remotes::install_github("matutosi/ensumr")
```

## Example

You can use dummy data.

```{r dummy_data}
library(ensumr)
dummy_data
```

You can show the summary of enquate by `enq_summary()`.

```{r enq_summary}
enq_summary(dummy_data)
enq_summary(dummy_data, simplify = TRUE)
```

You can analyse a item by `enq_simple()`, which can be used for both singular or multiple options. 
When the column name has "_ord", the result will be arranged by options. 
Otherwise, the results will be arranged by counted numbers.
When multiple options will be separated by deliminator (default is ";"). 

```{r enq_simple}
enq_simple(dummy_data, "q1")
enq_simple(dummy_data, "q2_ord")
enq_simple(dummy_data, "q3_multi")
```

You can analyse two item (cross tabulation) by `enq_cross()`. 

```{r enq_cross}
enq_cross(dummy_data, "q1", "q2_ord")
enq_cross(dummy_data, "q1", "q3_multi")
enq_cross(dummy_data, "q2_ord", "q3_multi", arrange_total = FALSE)
enq_cross_split(dummy_data, "q1", "q2_ord", "q3_multi")
```

## Citation

Toshikazu Matsumura (2024) ensumr. Analyse Enquate Data Easily. https://github.com/matutosi/ensumr/.
